\chapter{Determine Scope}

\textbf{Purpose:} The ontology models a modern streaming platform ecosystem. It integrates production data (creators, release dates) with a robust content catalog (Films, Series, Episodes, Age Ratings) and consumer access data (streaming platforms, subscriptions, viewing history) to support intelligent content recommendation and access querying.

\vspace{0.5cm}
\noindent\textbf{Competency Questions (SPARQL):}
\begin{enumerate}
    \item What are all the films in the catalog?
    \item What are the names of all the subscribers?
    \item Which streaming platforms are available in Romania?
    \item What is the URL of each streaming platform?
    \item Which subscription plans cost less than \$10?
    \item Which artworks belong to the Science Fiction genre?
    \item What plans does Netflix offer?
    \item Which subscribers have watched artworks that their friends also liked?
    \item Which subscribers have watched at least one science fiction film?
    \item Which subscribers are subscribed to a platform that is not available in their country?
    \item What are all reviews with their score and the name of the reviewed artwork?
    \item Who are the five oldest subscribers?
    \item Which subscribers liked an artwork but never wrote a review for it?
    \item What are all the episodes of \textit{Better Call Saul} Season 1?
    \item For each subscriber, what plan are they on and what does it cost — including subscribers with no plan?
    \item How many artworks has each subscriber watched, sorted from most to least?
    \item What is the average review score per genre, for genres with at least two reviews?
    \item Starting from the first episode of \textit{Better Call Saul} Season 1, what is the full episode sequence?
    \item Which pairs of subscribers are friends and have both watched the same artwork?
    \item For each platform, what is the cheapest, most expensive, and average plan cost?
    \item Which subscribers are older than 25?
    \item Which reviews have a score of 7 or above?
    \item Which subscription plans have no advertisements?
    \item If a subscriber has watched a series, which seasons have they also implicitly watched?
    \item Which artworks should be recommended to a subscriber based on what their friends like?
    \item Which artworks should be recommended to a subscriber based on shared genre preferences?
    \item Which subscribers have both a premium plan and are located in a country where their platform is available?
    \item Which films should be recommended to a subscriber because a friend likes the same genre and the film has a score of at least 9?
    \item Which pairs of subscribers are friends who have watched the same artwork?
    \item For a subscriber who has watched at least one episode in a season, which other episodes in that season are considered in progress? 
\end{enumerate}
